A Constraint Solving Approach to Parikh Images of Regular Languages

### 2024-03-01 calculus.tex
Ensure that figures are refered to, come before ref.
Closes: #78
Closes: #87
Fix warnings

We use the convention of splitting the formulas in proof goals into
linear (in)equalities ($\SomeInequalities$) and other formulas
($\SomeClause$), and assume that predicates $\Image$ and $\Connected$
only occur positively.
%Since our rules operate by adding and matching
%linear (in)equalities in a proof goal; we use the shorthand of listing
%the matched inequalities as antecedents, e.g., in \Propagate{}.
%
The transition selection function~$\Filter$ is represented
symbolically and can, in practice, be read as a function from
transitions to $\Naturals$-valued terms (e.g.~\texttt{t}
or~\texttt{t+1}). In our implementation \Catra{}, described in
\cref{sec:implementation}, $\Filter$~is a vector of fresh variables
with the same size as~$\Transitions$.

@@

\medskip
%
The rule~\textsc{Expand} expands an $\SinglePredicateInstance$
predicate into the more basic predicate
$\Connected(\Automaton, \Filter)$, as well as linear equations
relating the transitions mentioned by~$\Filter$ with the monoid
element~$\MonoidElement$, and linear flow equations described
by~\FlowEq{} (below). Since $\Connected$ and $\Image$ are partially
redundant and the difference is covered by~$\FlowEq$, we can remove
the instance of~$\Image$ when applying~$\Expand$. In this sense, we
split the semantics of the $\Image$~predicate into its counting aspect
(covered by $\FlowEq$) and its connectedness aspect (covered by
$\Connected$).

@@

\medskip
%
A decision procedure would start from one or multiple
predicates~$\SinglePredicateInstance$ to be satisfied, possibly in
combination with other constraints about $m$. It would then first
expand the predicates using the $\Expand{}$~rule, and subsequently
apply the other rules to search for a solution.

@@

The proof tree is shown in \cref{fig:derivation:single}.  Like in
\cref{sec:a_1,sec:intuition:algebra}, by arithmetic reasoning the
calculus ends up with fixed values for several of the transition
variables $\TransitionVar_k$, and eventually concludes that the
transition directly from state $I_{A}$ to $F_{A}$ (variable~$x_3$)
is incompatible with
the constraint.  It is then possible to remove the $\Connected{}$
predicate using \Subsume{}, since \Propagate{} is not able to infer
further constraints, and no non-trivial applications of \Split{}
remain to be done.  After this, by further arithmetic reasoning we can
derive a solution~$l_A = 4, i = 1, n = 2$, and conclude that the root
constraint is satisfiable.  To obtain values for the string
variables~$s_1, s_2$ corresponding to the solution, one can
construct an accepting path of the automaton with each transition~$t$
taken $\Filter(t)$ times.

\begin{figure}[tb]

@@

@@

>>>>

We use the convention of splitting the formulas in proof goals into
linear (in-) equalities ($\SomeInequalities$) and other formulas
($\SomeClause$), and assume that predicates $\Image$ and $\Connected$
only occur positively.
%Since our rules operate by adding and matching
%linear (in)equalities in a proof goal; we use the shorthand of listing
%the matched inequalities as antecedents, e.g., in \Propagate{}.
%
The transition selection function~$\Filter$ is represented
symbolically and can, in practice, be read as a function from
transitions to $\Naturals$-valued terms (e.g.~\texttt{t}
or~\texttt{t+1}). In our implementation \Catra{}, described 
in~\cref{sec:implementation}, $\Filter$~is a vector of fresh variables
with the same size as~$\Transitions$.

@@

The rule~\textsc{Expand} expands an $\SinglePredicateInstance$
predicate into the more basic predicate
$\Connected(\Automaton, \Filter)$, as well as linear equations
relating the transitions mentioned by~$\Filter$ with the monoid
element~$\MonoidElement$, and linear flow equations described
by~\FlowEq{} (below). Since $\Connected$ and $\Image$ are partially
redundant and the difference is covered by~$\FlowEq$, we can remove
the instance of~$\Image$ when applying~$\Expand$. In this sense, we
split the semantics of the $\Image$~predicate into its counting aspect
(covered by $\FlowEq$) and its connectedness aspect (covered by
$\Connected$).

@@

A decision procedure would start from one or multiple
predicates~$\SinglePredicateInstance$ to be satisfied, possibly in
combination with other constraints about $m$. It would then first
expand the predicates using the $\Expand{}$~rule, and subsequently
apply the other rules to search for a solution.

@@

\begin{figure}[h]

@@

\Description[]{}% This is fine; it's all text!

@@

The proof tree is shown in~\cref{fig:derivation:single}. Like in
\cref{sec:a_1,sec:intuition:algebra}, by arithmetic reasoning the
calculus ends up with fixed values for several of the transition
variables $\TransitionVar_k$, and eventually concludes that the
transition directly from state $I_{A}$ to $F_{A}$ (variable~$x_3$)
is incompatible with
the constraint.  It is then possible to remove the $\Connected{}$
predicate using \Subsume{}, since \Propagate{} is not able to infer
further constraints, and no non-trivial applications of \Split{}
remain to be done.  After this, by further arithmetic reasoning we can
derive a solution~$l_A = 4, i = 1, n = 2$, and conclude that the root
constraint is satisfiable.  To obtain values for the string
variables~$s_1, s_2$ corresponding to the solution, one can
construct an accepting path of the automaton with each transition~$t$
taken $\Filter(t)$ times.
